To log or not to log

### 2022-08-30 analysis/Analysis ECDC forecasts.Rmd
update paper draft

```{r}
# "CovidMetrics-epiBATS" only has forecasts in DE
# "ICM-agentModel" only in Poland
# "UVA is everywhere

plot_pred_score <- function(hub_data, scores, model,
                            horizon = 2, type = "Cases", 
                            remove_x_text = TRUE) {
  forecast_model <- model
  h <- horizon
  filtered_hub <- hub_data |>
    filter(model == forecast_model, 
           horizon %in% h, 
           target_type == type)

@@

# ICM <- plot_pred_score(hub_data, scores, "ICM-agentModel")
# epiBATS <- plot_pred_score(hub_data, scores, "CovidMetrics-epiBATS")
# BSLCoV <- plot_pred_score(hub_data, scores, "UB-BSLCoV")
# ggsave(filename = "output/figures/HUB-model-ICM.png", width = 9, height = 5, plot = ICM)
# ggsave(filename = "output/figures/HUB-model-epiBATS.png", width = 9, height = 5, plot = epiBATS)

ensemble_cases <- plot_pred_score(hub_data |> filter(location == "DE"), 
                                  scores, "EuroCOVIDhub-ensemble", 
                                  remove_x_text = TRUE, 
                                  type = "Cases")

ensemble_deaths <- plot_pred_score(hub_data |> filter(location == "DE"), 
                                   scores, "EuroCOVIDhub-ensemble", 
                                   type = "Deaths")

put_plot_together <- function(plot_cases, 
                              plot_deaths) {
  layout <- "
AE
BF
CG
DH"

@@

put_plot_together(ensemble_cases, ensemble_deaths)
ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 9, height = 7)

ensemble_cases <- plot_pred_score(hub_data |> filter(location == "DE"), 
                                  scores, "epiforecasts-EpiNow2", 
                                  remove_x_text = TRUE, 
                                  type = "Cases")

ensemble_deaths <- plot_pred_score(hub_data |> filter(location == "DE"), 
                                   scores, "EuroCOVIDhub-baseline", 
                                   type = "Deaths")

# baseline <- plot_pred_score(hub_data |> filter(location == "DE"), scores, "EuroCOVIDhub-baseline") 
# epinow <- plot_pred_score(hub_data |> filter(location == "DE"), scores, "epiforecasts-EpiNow2") 
# 
# ensemble / epinow +
#   plot_layout(guides = "collect", 
#               heights = c(1, 1, 2)) +
#   theme(legend.position = "bottom")
# 
# ggsave(filename = "output/figures/HUB-model-comparison.png", width = 9, height = 7)

>>>>

```{r}
plot_pred_score <- function(hub_data, scores, model,
                            horizon = 2, type = "Cases", 
                            remove_x_text = TRUE) {
  forecast_model <- model
  h <- horizon
  filtered_hub <- hub_data |>
    filter(model == forecast_model, 
           horizon %in% h, 
           target_type == type)

@@

put_plot_together <- function(modelname = "EuroCOVIDhub-ensemble", 
                              locationname = "DE") {

  plot_cases <- plot_pred_score(hub_data |> filter(location == locationname), 
                                  scores, modelname, 
                                  remove_x_text = TRUE, 
                                  type = "Cases")

  plot_deaths <- plot_pred_score(hub_data |> filter(location == locationname), 
                                     scores, modelname, 
                                     type = "Deaths")

  layout <- "
AE
BF
CG
DH"

@@

put_plot_together("EuroCOVIDhub-ensemble")
ggsave(filename = "output/figures/HUB-model-comparison-ensemble.png", width = 9, height = 7)

put_plot_together("epiforecasts-EpiNow2")
ggsave(filename = "output/figures/HUB-model-comparison-epinow.png", width = 9, height = 7)

put_plot_together("EuroCOVIDhub-baseline", locationname = "DE")
ggsave(filename = "output/figures/HUB-model-comparison-baseline.png", width = 9, height = 7)

### 2022-08-30 main.tex
update paper draft

% SECTION STRUCUTRE / IMPORTANT THINGS
% \begin{itemize}
% \item X variation gets smaller / more even
% \item X relationship with country size changes
% \item X inndividual rankings stay similar
% \item X relative skill scores change a lot
% \item deep dive into what happens for individual models
% \item scores increase across forecast horizons
% \end{itemize}
% 
% Maybe look at correlation between bias and score?
% Correlation between relative WIS contributions.?

@@

When comparing examples of forecasts on the natural scale with log-transformed ones (see Figures XX, XX, XX) a few interesting patterns emerge. Missing the peak, i.e. predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the log scale (see forecasts in May in Figure XXA, B). On the log scale, these have less of an influence, as errors are smaller in relative terms (see panels C and D). We hypothesise that the larger divergence in scores for cases (compared to scores for deaths) displayed in Figure \ref{fig:HUB-cors}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths. 
Conversely, failure to predict an upswing while numbers are still low, is not severely punished on the natural scale (see forecasts in July in Figure XX A, B), as overall absolute errors are low. On the log scale, missing lower inflection points tends to lead to more severe penalties (see panels C and D). One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure XXE, F). On the log scale, scores increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations. Furthermore, including very low numbers in the confidence intervals may lead to very high scores on the log-scale. This may be an issue especially when forecasting small quantities. Figure XXX shows forecasts and scores for the baseline model, which had zero included in its 50 percent intervals for the forecasts in XX to XX, leading to excessive dispersion values on the log scale. Notwithstanding the general concern, one can argue that in the specific example including zero in the prediction intervals was unreasonable and rightly penalised given that case numbers were over 1000.

@@

\subsection{Scoring the growth rate directly}
By dividing every observed value and every forecast by the last observed value prior to evaluation, we replace forecasting $y_{t+1}$ with forecasting  $y_{t+1} / y_t = 1 + g_{t, t+1}$. Just as we were previously penalised an additive error with respect the absolute value, we are now penalising an absolute error with respect to the growth rate.

We can illustrate what's happening by thinking of an analogous point prediction $\hat{g_{t, t+1}}$ for the growth rate. We could then write
\begin{equation}
    1 + g_{t, t+1} = 1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}. 
\end{equation}
%
The WIS and CRPS would evaluate the forecast based on the error $\varepsilon^*_{t+1}$. This translates to an absolute error on the actual forecast target that is multiplicatively connected to the last observed value: 
\begin{align}
    y_{t+1} &= (1 + \hat{g}_{t, t+1} + \varepsilon^*_{t+1}) \cdot y_t \\
            &= (1 + \hat{g}_{t, t+1}) \cdot y_t + \varepsilon^*_{t+1} \cdot y_t \\
            &= \hat{y}_{t+1} + \varepsilon^*_{t+1} \cdot y_t.
\end{align}

One disadvantage of this is that we can't really do anything when an observation is zero. 
%is this helpful?%
%Maybe add some math on how we actually transform the distribution. E.g. divide every quantile or sample by the last observed value, divide the distribution?%

\subsection{Changes in the WIS for forecasts on the log scale} \label{wis-log-derivation}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/placeholder-image.png}
    \caption{Placeholder}
    \label{fig:placeholder}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-mean-scores-vs-total.png}
    \caption{Plot with Weighted interval scores against the mean number of observed cases or deaths.}
    \label{fig:HUB-mean-scores-total}
\end{figure}

@@

%\section{Modelling scores}
%\paragraph{Motivation} Alternative is to use summary measures or pairwise comparison, which becomes cumbersome for many dimensions.

%\paragraph{caveats, practical limitations}
%\begin{itemize}
%    \item What kind of transformations can we do? Propose random effect models
%    \item is there anything we can't do? 
%    \item How does logging change the error distribution? On the natural scale, you have huge outliers. What's the appropriate error distribution to use? confidence intervals. Maybe we can only say something about the effects
%\end{itemize}

%Also: what, empirically is the distribution of errors?

>>>>

% SECTION STRUCUTRE / IMPORTANT THINGS
% \begin{itemize}
% \item X variation gets smaller / more even
% \item X relationship with country size changes
% \item X inndividual rankings stay similar
% \item X relative skill scores change a lot
% \item deep dive into what happens for individual models
% \item scores increase across forecast horizons
% \end{itemize}
% 
% Maybe look at correlation between bias and score?
% Correlation between relative WIS contributions on the log vs. the natural scale.?

@@

When comparing examples of forecasts on the natural scale with log-transformed ones (see Figures XX, XX, XX) a few interesting patterns emerge. Missing the peak, i.e. predicting increasing numbers while actual observations are already falling, tends to contribute a lot to overall scores on the log scale (see forecasts in May in Figure XXA, B). On the log scale, these have less of an influence, as errors are smaller in relative terms (see panels C and D). We hypothesise that the larger divergence in scores for cases (compared to scores for deaths) displayed in Figure \ref{fig:HUB-cors}C is related to the fact that outlier forecasts missing the peak are more common for cases than for deaths. 
Conversely, failure to predict an upswing while numbers are still low, is not severely punished on the natural scale (see forecasts in July in Figure XX A, B), as overall absolute errors are low. On the log scale, missing lower inflection points tends to lead to more severe penalties (see panels C and D). One can also observe that on the natural scale, scores tend to track the overall level of the target quantity (compare for example forecasts for March-July with forecasts for September-October in Figure XXE, F). On the log scale, scores increase whenever forecasts are far away from the truth in relative terms, regardless of the overall level of observations. Furthermore, including very low numbers in the confidence intervals may lead to very high scores on the log-scale. This may be an issue especially when forecasting small quantities. Figure \ref{fig:HUB-model-comparison-baseline} shows forecasts and scores for the baseline model, which had zero included in its 50 percent intervals for the forecasts in XX to XX, leading to excessive dispersion values on the log scale. Notwithstanding the general concern, one can argue that in the specific example (at least for cases) including zero in the prediction intervals was unreasonable and rightly penalised given that case numbers were in the thousands.

@@

\subsection{PROBABLY DELETE? Changes in the WIS for forecasts on the log scale} \label{wis-log-derivation}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-baseline.png}
    \caption{
    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany. A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale. B, F: Corresponding scores. C, G: Forecasts and observations on the log scale. D, H: Corresponding scores. 
    }
    \label{fig:HUB-model-comparison-baseline}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{output/figures/HUB-model-comparison-epinow.png}
    \caption{
    Forecasts and scores for two-week-ahead predictions from the EuroCOVIDhub-ensemble made in Germany. A, E: 50\% and 90\% prediction intervals and observed values for cases and deaths on the natural scale. B, F: Corresponding scores. C, G: Forecasts and observations on the log scale. D, H: Corresponding scores. 
    }
    \label{fig:HUB-model-comparison-epinow}
\end{figure}

@@
